The impact of gas bulk rotation on the Lyman-$\alpha$ line

### 2014-01-09 paper/ms.tex
minor updates

To quantify the line broadening we use the width at half maximum
(FWHM), $W$. Figure \ref{fig:widthvsvelocity} shows how $W$ increases
with rotational velocity. Continuous (dashed) lines connect the
results for homogeneous (central) source distribution. For the
temperature $T=10^4$K used in our radiative transfer calculations the
thermal velocity is $v_{th}=12.8$\kms. For a model with $\tau_{H}$ it
means that the FWHM increase up to $350$\kms (at $V_{\rm
  max}=300$\kms) compared to $150$\kms in the static case.

@@

The variations in the $\mu$ distribution is shown in Fig.\ref{fig:muhisto},
for the Off-Center model there is a strong dependence of the 
observed flux on the relative position of an observer with $\Delta\mu\leq 25\%$.
In contrast,the central and homogeneous source distributions only show a 
weak anisotropy at the level of $\Delta\mu\leq 3\%$. On the other hand the variations
in $\mu$ does not depend on the rotational velocity while increasing the optical depth $\tau_{H}$
would also increase the variation of $\Delta\mu$ up to a $15\%$.

Figure \ref{fig:OC_profiles} show the spectra for the two off-centred
spheres located at $C=\pm0.5R$.  In this case we only use photons that
have outgoing directions aligned with the observer ($\vert \mu \vert > 
0.9$ $\&$ $\vert \mu \vert < 1.0]$).  We do not find any asymmetry in 
the line produced by the off-center emission, The difference
 between the profiles corresponding to the spheres located at $C=-0.5R$ 
 and that ones located at $C=0.5R$ is negligible.  As in the homogeneous / 
central model the line gets broader as the rotational velocity increase also 
the central part of the line becomes predominant . 
There is also a signature of possible three peaks profiles for some models ($\tau = 10^6,
V=100,200$\kms and $\tau = 10^7, V=200,300$\kms) which would be studied 
in detail in \S \ref{sec:3p}.

@@

As mentioned in \S \ref{sec:off-center} Off-centered emission shows signs of three peak 
profiles, to analyse this in more detail we use the same method of the spheres
explained in \S \ref{sec:off-center}. We select spheres at $C=(\pm 0.25R, \pm0.5R, \pm0.75R)$ with a radii of
$0.25R$ for all the models, for those who present signs of three peak profiles such as 
$(V_{max}=50kms^{1}$ \& $\tau_{H}=10^{5}), (V_{max}=100kms^{1}$ \& $\tau_{H}=10^{6}), 
(V_{max}=200kms^{1}$ \& $\tau_{H}=10^{7})$ we made another run with $\tau_{\gamma}=10^{6}$ 
with the aim of get a stronger profile.

Fig. \ref{fig:3p_profiles} shows our findings, in order to get such profiles 
a relation between optical depth, rotational velocity and $C$ must be accomplish.
In particular we found that for the outer spheres $C=\pm 0.75$ the three
peak arises for all of our models with $\tau_{\gamma}=10^{6}$ this is due to 
the fact that the wings of the line move away from the central part as $C$ increase. 
Another requirement is that a high optical depth must be balanced with a high velocity 
in order to make the wings and the central peak similar in 
height.

@@

\item For homogeneous sources we find the changes in the
line morphology are linked to the reduction in the average number of
scatterings. In the second part of \S \ref{sec:widthpeak} we show
that, as rotational velocity increases, a large fraction of photons
escape with less than $10$ scatterings and on average 
with $40\%$ less scatterings than in the static case. This makes highly improbable 
that a photon's frequency can move away from the line's center. 
In \S \ref{sec:widthpeak} For typical column densities of $N_{HI}=1^{19}-1^{20}$, 
we find that rotational velocities of 100-300 \kms broaden the line profile up 
to a factor of $2-3$ in comparison to the static case. Simultaneously, the
flux in the line center also increases with rotational velocity. This
increase induces the transition from a double peak to a single peak
line in the models where the radiation sources are homogeneously
distributed. Also the escape fraction increase by factors of $2-10$ with 
respect to the slab with centrally sources.

\item For non-symmetric sources we find that three peak profiles
could emerge for models in which the rotational velocity is the necessary 
to give rise to a central peak but also not that high to decrease the
amount of photons in the wings. When initial off-center sources are 
selected from the outer parts of the galaxy the wings of the line are stronger.
Also off-center emission do not produce any considerable asymmetries in the line.\\
\end{itemize}

>>>>

To quantify the line broadening we use the width at half maximum
(FWHM), $W_{1/2}$. Figure \ref{fig:widthvsvelocity} shows how $W_{1/2}$ increases
with rotational velocity. Continuous (dashed) lines connect the
results for homogeneous (central) source distribution. For the
temperature $T=10^4$K used in our radiative transfer calculations the
thermal velocity is $v_{th}=12.8$\kms. For a model with $\tau_{H}$ it
means that the FWHM increase up to $350$\kms (at $V_{\rm
  max}=300$\kms) compared to $150$\kms in the static case.

@@

The variations in the $\mu$ distribution is shown in Fig.\ref{fig:muhisto},
for the Off-Center model there is a strong dependence of the 
observed flux on the relative position of an observer with $\Delta\mu\leq 25\%$.
In contrast,the central and homogeneous source distributions only show a 
weak anisotropy at the level of $\Delta\mu\leq 3\%$. The variations
in $\mu$ does not depend on the rotational velocity while increasing the optical depth $\tau_{H}$
would also increase the variation of $\Delta\mu$ up to a $15\%$.

Figure \ref{fig:OC_profiles} show the spectra for the two off-centered
spheres located at $C=\pm0.5R$.  In this case we only use photons that
have outgoing directions aligned with the observer i.e $\vert \mu \vert > 
0.9$ $\&$ $\vert \mu \vert < 1.0]$.  We do not find any asymmetry in 
the line produced by the off-center emission, the difference
 between the profiles corresponding to the spheres located at $C=-0.5R$ 
 and that ones located at $C=0.5R$ is negligible.  As in the homogeneous / 
central model the line gets broader as the rotational velocity increase also 
the central part of the line becomes predominant . 
There is also a signature of possible three peaks profiles for some models ($\tau = 10^6,
V=100,200$\kms and $\tau = 10^7, V=200,300$\kms) which would be studied 
in detail in the next section.

@@

As mentioned above Off-centered emission shows signs of three peak 
profiles, to analyse this in more detail we use the same method of the spheres
explained in \S \ref{sec:off-center}. We select spheres at $C=(\pm 0.25R, \pm0.5R, \pm0.75R)$ with a radii of
$0.25R$ for all the models, for those who present signs of three peak profiles such as 
$(V_{max}=50km$ $s^{-1}$ \& $\tau_{H}=10^{5}), (V_{max}=100km$ $s^{-1}$ \& $\tau_{H}=10^{6}), 
(V_{max}=200km$ $s^{-1}$ \& $\tau_{H}=10^{7})$ we made another run with $\tau_{\gamma}=10^{6}$ 
with the aim of get a stronger profile.

Fig. \ref{fig:3p_profiles} shows our findings, in order to get such profiles 
a relation between optical depth, rotational velocity and $C$ must be accomplish.
In particular we found that for the outer spheres $C=\pm 0.75$ the three
peak arises for all of our models with $\tau_{\gamma}=10^{6}$ this is due to 
the fact that the wings of the line move away from the central part as $C$ increase. 
Another requirement is that a high optical depth must be balanced with a high velocity 
in order to make the wings and the central peak similar in 
height, for models in which the velocity is low spheres at $C = \pm 0.25$ produce 
noise three peak profiles.

@@

\item For homogeneous sources we find the changes in the
line morphology are linked to the reduction in the average number of
scatterings. In the second part of \S \ref{sec:widthpeak} we show
that, as rotational velocity increases, a large fraction of photons
escape with less than $10$ scatterings and on average 
with $40\%$ less scatterings than in the static case. This makes highly improbable 
that a photon's frequency can move away from the line's center. 
In \S \ref{sec:widthpeak} For typical column densities of $N_{HI}=1e^{19}-1e^{20}$, 
we find that rotational velocities of 100-300 \kms broaden the line profile up 
to a factor of $2-3$ in comparison to the static case. Simultaneously, the
flux in the line center also increases with rotational velocity. This
increase induces the transition from a double peak to a single peak
line in the models where the radiation sources are homogeneously
distributed. Also the escape fraction increase by factors of $2-10$ with 
respect to the slab with centrally sources.

\item For non-symmetric sources we find that three peak profiles
could emerge for models in which the rotational velocity is the necessary 
to give rise to a central peak but also not that high to decrease the
amount of photons in the wings. When initial off-center sources are 
selected from the outer parts of the galaxy the wings of the line are stronger.
Also off-center emission do not produce any considerable asymmetries in the line.
\end{itemize}
